Deriving Hamiltonian mechanics from physics, without passing through Lagrangians.

### 2014-06-15 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
More notation and language cleanup

\begin{prop}\label{detMap}
Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:i \in I \leftrightarrow i' \in I$.
\end{prop}

@@

\begin{defn}\label{labelsIndep}
Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$. Let $f:i \in I \leftrightarrow i' \in I$ be a bijective map. The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate mappings:
\begin{enumerate}
\item $I_1' \times I_2' = I$
\item there exists a bijective map $f_1: i_1 \in I_1 \leftrightarrow i_1' \in I_1'$
\item there exists a bijective map $f_2: i_2 \in I_2 \leftrightarrow i_2' \in I_2'$
\item $f(i) = \{f_1(i_1), f_2(i_2)\}$
\end{enumerate}
\end{defn}

@@

\begin{cor}\label{discreteEv}
Let $f: i \in I \leftrightarrow i'$ be a bijective map, we have $\mathbbm{c}_j=\sum D'(i') \mathbbm{s}_i'=\sum \rho'(i') w(i') \mathbbm{s}_i' = \sum D(i) \mathbbm{s}_i' = \sum \rho(i) w(i) \mathbbm{s}_i'$. $\rho'(i') = \rho(i) w(i) / w(i')$.
\end{cor}

@@

With homogeneous labels no adjusting is needed, and the range can be used as a measure of the number of labels.

@@

\begin{prop}\label{continuousMapping}
Let $f: X \mapsto X$ be a bijective map on a state variable. The mapping must be continuous.
\end{prop}
%http://en.wikipedia.org/wiki/Limiting_density_of_discrete_points
Assume mapping is discontinuous at point $x$. Consider the cell at $x$ of width $m(x)dx$. The cell would be split into two, so it would not be mapped to one and only one cell.

\begin{prop}\label{widthMapping}
Let $f: X \mapsto X$ be a deterministic and reversible mapping on a state variable. Then $m(x') dx' \equiv m(f(x)) \frac{\partial f(x)}{\partial x} dx = m(x) dx$. If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the number of states.
\end{prop}
The mapping must be done so that the width of the cells is mapped as well, not just the center value.

@@

\begin{prop}\label{sdofMap}
Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
\end{prop}

@@

Simply applying the vector transformation rules under continuous transformation we have:
\begin{align*}
V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} &= V'^{\alpha} \Omega_{\alpha, \beta} W'^{\beta}  \\
&= (V^{\alpha} + \partial_{\gamma} S^{\alpha} dt V^{\gamma}) \Omega_{\alpha, \beta} ( W^{\beta} + \partial_{\delta} S^{\beta} W^{\delta} dt) \\
&= V^{\alpha} \Omega_{\alpha, \beta} W^{\beta} + (\partial_{\gamma} S^{\alpha} V^{\gamma} \Omega_{\alpha, \beta} W^{\beta} \\
 &+ V^{\alpha} \Omega_{\alpha, \beta} \partial_{\delta} S^{\beta} W^{\delta}) dt + O(dt^2)
\end{align*}
\begin{align*}
V^{\gamma} W^{\beta} \partial_{\gamma} S_{\beta} - V^{\alpha} W^{\delta} \partial_{\delta} S_{\alpha} = 0
\end{align*}
\begin{align*}
\partial_{\alpha} S_{\beta} - \partial_{\beta} S_{\alpha} &= curl(S_{\alpha}) = 0 \\
S_{\alpha} &= \partial_{\alpha}H
\end{align*}

@@

The trajectory has to be continuous in both spatial and temporal variables because of \ref{continuousMapping}. The mapping still needs to give one and only one value in time, so we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$. Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional. What is physical, instead, is that $dt/ds$ along a trajectory cannot change sign, so particle and anti-particle states cannot be connected by deterministic and reversible evolution.

>>>>

\begin{prop}\label{detMap}
Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
\end{prop}

@@

\begin{defn}\label{labelsIndep}
Let $I_1$ and $I_2$ be two orthogonal sets of labels and $I = I_1 \times I_2$. Let $f:I \leftrightarrow I$ be a bijective map. The labels are said independent iff we can define two sets of orthogonal labels $I_1'$ and $I_2'$ such that we can define separate mappings:
\begin{enumerate}
\item $I_1' \times I_2' = I$
\item there exists a bijective map $f_1: I_1 \leftrightarrow I_1'$
\item there exists a bijective map $f_2: I_2 \leftrightarrow I_2'$
\item $f(i) = \{f_1(i_1), f_2(i_2)\}$
\end{enumerate}
\end{defn}

@@

\begin{cor}\label{discreteEv}
Let $f: I \leftrightarrow I$ be a bijective map, we have $\mathbbm{c}_j=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
\end{cor}

@@

With homogeneous labels no adjustment is needed, and the range can be used as a measure of the number of labels.

@@

\begin{prop}\label{continuousMapping}
Let $f: X \leftrightarrow X$ be a bijective map on a state variable. The mapping must be continuous.
\end{prop}
%http://en.wikipedia.org/wiki/Limiting_density_of_discrete_points
Assume mapping is discontinuous at point $x$. Consider the cell at $x$ of width $m(x)dx$. The cell would be split into two, so it would not be mapped to one and only one cell.

\begin{prop}\label{widthMapping}
Let $f: X \leftrightarrow X$ be a bijective map on a state variable. Then $m(x') dx' \equiv m(f(x)) \frac{\partial f(x)}{\partial x} dx = m(x) dx$. If $X$ is homogeneous, then $dx' = dx$ and the range gives us a measure of the number of states.
\end{prop}
The mapping must be done so that the width of the cells is mapped as well, not just the center value.

@@

\begin{prop}\label{sdofMap}
Let $f: X \leftrightarrow X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
\end{prop}

@@

Simply applying the vector transformation rules under continuous transformation we have:
\begin{align*}
v^{\alpha} \omega_{\alpha, \beta} w^{\beta} &= v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}  \\
&= (v^{\alpha} + \partial_{\gamma} S^{\alpha} dt v^{\gamma}) \omega_{\alpha, \beta} ( w^{\beta} + \partial_{\delta} S^{\beta} w^{\delta} dt) \\
&= v^{\alpha} \omega_{\alpha, \beta} w^{\beta} + (\partial_{\gamma} S^{\alpha} v^{\gamma} \omega_{\alpha, \beta} w^{\beta} \\
 &+ v^{\alpha} \omega_{\alpha, \beta} \partial_{\delta} S^{\beta} w^{\delta}) dt + O(dt^2)
\end{align*}
\begin{align*}
v^{\gamma} w^{\beta} \partial_{\gamma} S_{\beta} - v^{\alpha} w^{\delta} \partial_{\delta} S_{\alpha} = 0
\end{align*}
\begin{align*}
\partial_{\alpha} S_{\beta} - \partial_{\beta} S_{\alpha} &= curl(S_{\alpha}) = 0 \\
S_{\alpha} &= \partial_{\alpha}H
\end{align*}

@@

The trajectory has to be continuous in both spatial and temporal variables because of \ref{continuousMapping}. Since determinism and reversibility are defined in time, the trajectory must traverse all times once and only once: we must have an invertible mapping between $t$ and $s$, which means we must have a strictly monotonic $t(s)$. $dt/ds$ along a trajectory cannot change sign, so we have the division between particle and anti-particle states. Note that since the parametrization is conventional and can be changed to $s'=-s$, what we call particle and anti-particle states is also conventional. What is physical and not conventional, though, is that particle and anti-particle states cannot be connected by deterministic and reversible evolution.
